---
title: "Ecography_Commented_Code"
author: "Adam Eichenwald, Sydne Record"
date: "2023-10-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(DescTools)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(lme4)
library(ggplot2)
library(ForestTools)
library(terra)
library(sf)
library(hexbin)
```


# Creating the canopy shapefile
``` {r}
Harvcanopy<- terra::rast("/Users/adameichenwald/Downloads/GLiHT_Prospect_Hill_Jun2012_CHM.tif")
```

# Plot CHM (extra optional arguments remove labels and tick marks from the plot)
```{r}
par(mar = rep(0.5, 4))
plot(Harvcanopy, xlab = "", ylab = "", xaxt='n', yaxt = 'n')
```
# Function for defining dynamic window size
```{r}
lin <- function(x){x * 0.05 + 0.6}
```
# Detect treetops
```{r}
ttops <- vwf(Harvcanopy, winFun = lin, minHeight = 2)
```
# Plot CHM
```{r}
plot(Harvcanopy, xlab = "", ylab = "", xaxt='n', yaxt = 'n')
```
# Add dominant treetops to the plot
```{r}
plot(ttops$geometry, col = "blue", pch = 20, cex = 0.5, add = TRUE)
```
# Create crown map
```{r}
crowns_ras <- mcws(treetops = ttops, CHM = Harvcanopy, minHeight = 1.5)
expanse(crowns_ras, unit = "m")
```

# Plot crowns
```{r}
plot(crowns_ras, col = sample(rainbow(50), nrow(unique(chm)), replace = TRUE), legend = FALSE, xlab = "", ylab = "", xaxt='n', yaxt = 'n')

class(crowns_ras)
crownspolygon<-as.polygons(crowns_ras)

writeVector(crownspolygon, "/Users/adameichenwald/Documents/University of Maine/treecrowns.shp")
```

# Data Import


```{r}
Harv2019 <- read_csv("Harv2019AllInformation.csv")
```

# 2019 models
Split the Harvard 2019 dataset into test and train data, so that we can test how well the model predictions work.
```{r}
sample <- sample(c(TRUE, FALSE), nrow(Harv2019),
                 replace=TRUE, prob=c(0.7,0.3))
train <- Harv2019[sample, ]
test <- Harv2019[!sample, ]                      

```

### Construct the models.
Note that we have included height and canopy diameter as interaction terms, using (*) rather than (+). This is to copy Jucker et al, who also use height and canopy diameter as interaction terms. 
```{r}
onlyhgtdiam<-lm(data = train,
                   log(dbh)~log(height*diameter))

includestage<-lm(data = train,
                         log(dbh)~log(height*
                           diameter)+
                           Stage)
includeshade<-lm(data = train,
                          log(dbh)~log(height*
                            diameter)+Shade)
includestageshade<-lm(data = train,
                          log(dbh)~log(height*
                            diameter)+
                            Stage+Shade)

list<-list(onlyhgtdiam,includeshade, includestage,includestageshade)
dataframe<-data.frame(model = c("onlyhgtdiam","includeshade",
                                "includestage","includestageshade"),
                      AIC = sapply(list, function(x) AIC = AIC(x)))%>%
  mutate(AIC = ifelse(model == "Harv2019mixedmodel", 
                      AIC(Harv2019mixedmodel),
                      AIC))%>%
  arrange(AIC)
dataframe

tab_model(includestageshade,includestage,includeshade,onlyhgtdiam,
          show.aicc = TRUE)

```



The best model according to AIC includes the effect of both shade and stage of succession as random variables (mixed model).

Test to see whether the variables are correlated.

``` {r}
chisq.test(Harv2019$Stage, Harv2019$Shade)
```
They are. Therefore, the second best model "stage only" is better, since including both variables violates regression assumptions.

### Test the best model
Predict test dbh from best model in previous section.
```{r}

predicted <- predict(includestage, test, type="response")

```

Predict test dbh using Jucker et al.'s allometric equation.
```{r}
Jucker<-test%>%
  #mutate(DBH_Jucker = 0.557 * 
   #        (height * diameter)^0.809 * 
  #         exp(0.056^2/2))%>%
#This is the same equation that is in Jucker et al. 2017, except transformed into logspace
mutate(DBH_Jucker = log(0.557) + 0.809 * log(height * diameter) + 
           ((0.056^2) / 2))
```

Examine the r2 and the slope of each model. We would expect the better of the two models to have slopes closer to 1 and have high r2 values.

```{r}
juckertestmodel<-lm(data=Jucker, log(dbh)~DBH_Jucker)
stagesmodel<-lm(log(test$dbh)~predicted)
summary(juckertestmodel)
summary(stagesmodel)
confint(juckertestmodel, level = 0.95)
confint(stagesmodel,  level = 0.95)
```

Our mixed model has a much higher r2 value and has a slope much closer to 1 than the predictions from Jucker et al, as we would expect based on our hypotheses.


```{r}

data.frame(Method = c("Jucker et al.", "Stage Mixed Model"),
           RMSE = c(RMSE(juckertestmodel), RMSE(stagesmodel)))


```
Our mixed model comes out with a lower RMSE, which again puts it on top as the better model.


### Plot the results
According to Pinero et al. 2008 (Ecological Modeling), "observed (in the y-axis) vs. predicted (in the x-axis) (OP) regressions should be used" rather than vice versa. This is because "a spurious effect is added to the regression parameters when regressing PO values and comparing them against the 1:1 line."

The line in the plots below is not the regression line. It is instead the 1:1 line.

```{r}
select<-dplyr::select
ggplotdata<-rbind(Jucker%>%
  select(DBH_Jucker, dbh, Stage, Shade)%>%
  rename(predicted = DBH_Jucker)%>%
  mutate(method = "Jucker et al. (2017)"),
test%>%
  select(dbh,Stage, Shade)%>%
  mutate(predicted = predicted)%>%
  mutate(method = "Mixed Model (Stage Only)"))
library(Rmisc)
ggplotdataCI<-ggplotdata%>%
  group_by(Stage, method)%>%
  dplyr::summarize(CI(predicted))%>%
  ungroup()%>%
  dplyr::mutate(variable=rep(c("high_x","mean_x","low_x"), 8))%>%
  pivot_wider(names_from="variable", values_from = "CI(predicted)")%>%
  cbind(ggplotdata%>%
          group_by(Stage, method)%>%
  dplyr::summarize(CI(dbh))%>%
  ungroup()%>%
  dplyr::mutate(variable=rep(c("high_y","mean_y","low_y"), 8))%>%
  pivot_wider(names_from="variable", values_from = "CI(dbh)")%>%
  select(-Stage, -method))%>%
  rbind(ggplotdata%>%
  group_by(Shade, method)%>%
  dplyr::summarize(CI(predicted))%>%
  ungroup()%>%
  dplyr::mutate(variable=rep(c("high_x","mean_x","low_x"), 6))%>%
  pivot_wider(names_from="variable", values_from = "CI(predicted)")%>%
  cbind(ggplotdata%>%
          group_by(Shade, method)%>%
  dplyr::summarize(CI(dbh))%>%
  ungroup()%>%
  dplyr::mutate(variable=rep(c("high_y","mean_y","low_y"),6))%>%
  pivot_wider(names_from="variable", values_from = "CI(dbh)")%>%
  select(-method, -Shade))%>%
  dplyr::rename(Stage = Shade))%>%
  dplyr::rename(Effect = Stage)%>%
  mutate(Effect= ifelse(Effect == "high tolerance",
                        "High shade", Effect))%>%
  mutate(Effect= ifelse(Effect == "intermediate tolerance",
                        "Intermediate\nshade", Effect))%>%
  mutate(Effect= ifelse(Effect == "low tolerance",
                        "Low shade", Effect))%>%
  mutate(Effect= ifelse(Effect == "all",
                        "All stages", Effect))%>%
   mutate(Effect= ifelse(Effect == "climax",
                        "Late stage", Effect))%>%
  mutate(Effect= ifelse(Effect == "early",
                        "Early stage", Effect))%>%
  mutate(Effect= ifelse(Effect == "mid",
                        "Mid stage", Effect))%>%
  mutate(Effect= factor(Effect, levels = c("Early stage",
                                           "Mid stage",
                                           "Late stage",
                                           "All stages",
                                           "Low shade",
                                            "Intermediate\nshade",
                                           "High shade")))%>%
  filter(Effect != "Low shade",
         Effect!= "Intermediate\nshade",
         Effect != "High shade")

CIfigure1<-ggplot(ggplotdataCI,
       aes(y=mean_y, x=mean_x, color = Effect))+geom_point()+
  geom_errorbarh(aes(xmax = high_x, xmin = low_x))+
  geom_errorbar(aes(ymax = high_y, ymin = low_y))+
  geom_abline(intercept = 0, slope = 1, lty = 2)+
  facet_wrap("method")+
  xlab("Predicted DBH")+ylab("Measured DBH")+
  theme_bw()+
  theme(text=element_text(size=13))+
  xlim(11,35)
#This is figure 4 in the paper
CIfigure1

```

```{r}
hex_scale_log_colors <- scale_fill_gradientn(colours = colorRampPalette(rev(RColorBrewer::brewer.pal(9, 'RdYlBu')), bias=1)(50),
                                             trans = 'log', name = 'Individuals',
                                            label = function(x) sprintf("%.2f", x))

heatplot1<-ggplot(ggplotdata, aes(y=dbh, x=predicted))+geom_hex()+
 hex_scale_log_colors+
  geom_abline(intercept = 0, slope = 1, lty = 2)+
  facet_wrap("method")+
  xlab("Predicted DBH")+ylab("Measured DBH")+
  theme_bw()+
  theme(text=element_text(size=13))

#This is figure 3 in the paper
heatplot1

```
